---
title: "Customer Rating of Breakfast Cereals"
output:
  html_document:
    df_print: paged
  html_notebook:
    highlight: textmate
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
dir.create("images")
```

***

# Problem Statement

The dataset Cereals.csv includes nutritional information, store display, and consumer ratings for 77 breakfast cereals.

* Data Preprocessing. Remove all cereals with missing values.
* Apply hierarchical clustering to the data using Euclidean distance to the normalized measurements. Use Agnes to compare the clustering from  single linkage, complete linkage, average linkage, and Ward. Choose the best method.
* How many clusters would you choose?
* Comment on the structure of the clusters and on their stability. Hint: To check stability,  partition the data and see how well clusters formed based on one part apply to the other part. To do this: 
* Cluster partition A
* Use the cluster centroids from A to assign each record in partition B (each record is assigned to the cluster with the closest centroid).
* Assess how consistent the cluster assignments are compared to the assignments based on all the data.
* The elementary public schools would like to choose a set of cereals to include in their daily cafeterias. Every day a different cereal is offered, but all cereals should support a healthy diet. For this goal, you are requested to find a cluster of "healthy cereals." Should the data be normalized? If not, how should they be used in the cluster analysis?

***

Some code in this package comes from the following reference:

* https://www.dataminingbook.com/book/r-edition

***

Install packages if necessary. Uncomment before running.

```{r}
library(caret)
library(cluster) # for Agnes

```

***

## Data Processing


```{r}
cereals.df <- read.csv("./Cereals.csv")
cereals.df.nm <- na.omit(cereals.df)
head(cereals.df.nm)

# Normalize

cereals.df.nm.norm <- scale(cereals.df.nm[,-c(1:3)])
summary(cereals.df.nm.norm)

```
## Apply Agnes

```{r}
hc_single <- agnes(cereals.df.nm.norm, metric="euclidean",method="single")
hc_complete <- agnes(cereals.df.nm.norm, metric="euclidean",method="complete")
hc_average <- agnes(cereals.df.nm.norm, metric="euclidean",method="average")
hc_ward <- agnes(cereals.df.nm.norm, metric="euclidean",method="ward")

# print values

print(hc_single$ac)
print(hc_complete$ac)
print(hc_average$ac)
print(hc_ward$ac)

```
Ward's is the best. We will use that going forward.

### Apply Clustering

```{r}
pltree(hc_ward, cex = 0.6, hang = -1, main = "Dendrogram of agnes")
rect.hclust(hc_ward, k = 5, border = 1:4)
```
A distance of about 10 appears to give us 5 clusters. 
```{r}

members <- cutree(hc_ward, k = 5) # cut 5 clusters

centers <- aggregate( . ~ members, data = cereals.df.nm[,-c(1:3)], FUN = mean)
centers

```

Cluster 5 is low in fat, sodium, and sugars, so probably represents the healthiest group of cereals.

#### Stability of Clusters

Let us now leave some observations off to determine cluster stability
```{r}
#remove around 10% (5% from head, 5% tail data) to check cluster stability
dim(cereals.df.nm.norm)
d.norm <- dist(cereals.df.nm.norm[-c(1:4, 71:74),], method = "euclidean")
hc1 <- hclust(d.norm, method = "ward")
plot(hc1, hang = -1, ann = FALSE)
```
While we still have 5 potential clusters, the arrangement appears to have shifted, thus signalling a lack of stability.