---
title: "MidTerm - KMeans"
author: "Murali Shanker"
date: "October 31, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem

The dataset on American College and University Rankings contains information on 1302 American colleges and universities offering an undergraduate program. For each university, there are 17 measurements, including continuous measurements (such as tuition and grad- uation rate) and categorical measurements (such as location by state and whether it is a private or public school).

Note that many records are missing some measurements. 

1.Remove all records with missing measurements from the dataset.
2.For all the continuous measurements, run K-Means clustering. Make sure to normalize the measurements. How many clusters seem reasonable for describing these data? What was your optimal K?
3.Compare the summary statistics for each cluster and describe each cluster in this context (e.g., “Universities with high tuition, low acceptance rate...”).
4.Use the categorical measurements that were not used in the analysis (State and Private/Public) to characterize the different clusters. Is there any relationship between the clusters and the categorical information?
5.What other external information can explain the contents of some or all of these clusters?
6.Consider Tufts University, which is missing some information. Compute the Euclidean distance of this record from each of the clusters that you found above (using only the measurements that you have). Which cluster is it closest to? Impute the missing values for Tufts by taking the average of the cluster on those measurements. 
***

##Read and Clean Data

```{r}
# load required packages
library(tidyverse)  # data manipulation
library(factoextra) # clustering algorithms & visualization
library(ISLR)
```

```{r}
# read data
univ <- read.csv("./Universities.csv")

# Standardize using all numeric data
univ.norm <- scale(univ[,-c(1:3)])
univ.norm <- cbind(univ[,1:3],univ.norm) # Public / Private is a categorical variable
univ.norm[,3] <- as.factor(univ.norm[,3])

# Extract the Tufts record
tuft.norm <- univ.norm[which(univ.norm$College.Name=="Tufts University"),]
tuft.norm

# Now remove all recods with missing values

univ.norm <- na.omit(univ.norm)
summary(univ.norm)
```
***
## Determine K
```{r}
fviz_nbclust(univ.norm[,-c(1:3)], kmeans, method = "wss")
fviz_nbclust(univ.norm[,-c(1:3)], kmeans, method = "silhouette")
```

It appears that the optimal number of clusters is 3. 

***
## Explore clusters
```{r}
set.seed(120)
k3 <- kmeans(univ.norm[,-c(1:3)], centers = 3, nstart = 25)
univ.k3center <- as.data.frame(k3$centers)

colnames(univ.k3center) <- c("Recd","Accpt","newSe","newS10","newS25","FTu","PTu","INt","OSt","room","board","fees","books","personal","FPhD","SFRatio","GRate")

# Add a column for cluster number / to be also used for color
cluster <- matrix(c("1","2","3"),nrow = 3)
univ.k3center <- cbind(cluster,univ.k3center)
```

We will use GGally to draw parallel plots
```{r}
library(hrbrthemes)
library(GGally)
library(viridis)

# We write two parallel plots for clarity
ggparcoord(univ.k3center,
    columns = 2:8, groupColumn = 1, 
    showPoints = TRUE, 
    title = "Parallel Coordinate Plot for University Data - K = 3",
    alphaLines = 0.3 
    ) 

ggparcoord(univ.k3center,
    columns = 9:18, groupColumn = 1, 
    showPoints = TRUE, 
    title = "Parallel Coordinate Plot for for University Data - K = 3",
    alphaLines = 0.3 
    ) 
```

### Observations

* The most number of applications received and accepted are in Cluster 1. But, Cluster 2 accepts the most high-quality students, i.e., new students from top 10 and top 25 schools.
* While tuition is high for Cluster 2, fees, books and personal income is low. The reverse is true for schools in Cluster 1. Cluster 3 is unremarkable in any aspect.
* Cluster 2 has the best values for PhD faculty, Student/Faculty ratio, and graduation rate. These schools appear to be most selective in their choice of students, have highly qualified faculty, and small class sizes. All factors conducive to high student graduation rates.
* While Cluster 1 has a high number of PhD faculty, their Student/Faculty ratios are high. Given the non-selective nature of admissions, poor Student/Faculty ratios, graduation rates are not high. This dispite the low tuition rates.
* Cluster 3 is unremarkable from any measure.

***

## Explore relationship

Here, we explore the relationship between the categorical variables and clusters
```{r}
library(ggplot2)

cls <- as.factor(k3$cluster) # Save the cluster values
ts <- cbind(cls,univ.norm[,1:3]) # combine with categorical data

ggplot(ts) +
 aes(x = State, fill = Public..1...Private..2.) +
 geom_bar() +
 scale_fill_hue() +
 theme_minimal() +
 facet_wrap(vars(cls))
```

The observations in this plot confirm our earlier findings. Cluster 1 contains mainly public universities. Such universities, like Kent State, have low tuition, but high related costs. They also have significant faculty with PhDs, but class sizes are large, and thus graduation rates low. Cluster 2 consists of private universities. These have high tuition costs, but well-qualified faculty and small class sizes. Graduation rates are thus also high.Cluster 3 conists of a mix of public and private universities, so unremarkable in any distiguishing way.

A breakdown by State or Type of Institution (4-year versus 2-year, PhD granting versus not, etc.) might provide finer assessments to the clusters.

***

## Prediction for Tufts

```{r}
tuft.dist <- rbind(k3$centers,tuft.norm[,-c(1:3)]) 
get_dist(tuft.dist)
```

Clearly, the shortest distance for Tufts Univesity is Cluster 2. That aligns with our discussion that Cluster 2 mainly consists of private universities.

Now, we replace the missing values in our Tufts record with the cluster centroids from Cluster 2
```{r}
cl2 <- as.list(k3$centers[2,])
library(tidyr)
tuft.norm.rep <- replace_na(tuft.norm,cl2)
tuft.norm.rep
```
