---
title: "Machine Learining_Project 1"
output: html_notebook
---

Read in file.Intall caret.
```{r}
UBdata<-read.csv("UniversalBank.csv")
head(UBdata)
summary(UBdata)

#install.packages("caret")
library(caret)
```

Transform categorical predictors with more than two categories into dummy variables. In this case, Education.
```{r}
UBdata$Education<-as.factor(UBdata$Education)
dummy_model<-dummyVars(~Education,data=UBdata)
head(predict(dummy_model, UBdata))
Education_dummy<-predict(dummy_model, UBdata)
Education_dummy
UBdata<-cbind(UBdata, predict(dummy_model, UBdata))
UBdata$Education<- NULL
UBdata
```

Nomralize Data
```{r}
Norm_UBdata<-preProcess(UBdata[, c(2, 3, 4, 6, 7, 8)], method = c("center", "scale"))
UBdata_normalized<- predict(Norm_UBdata,UBdata)
summary(UBdata_normalized)
```

Partition the data into training (60%) and validation (40%) sets.
```{r}
Train_Index<-createDataPartition(UBdata$Personal.Loan, p = 0.6, list=FALSE)
Train_Data<-UBdata[Train_Index,]
Test_Data<-UBdata[-Train_Index,]
```

Examine that that 2 sets have similar distribution in personal loan acceptance rate
```{r}
summary(Train_Data$Personal.Loan)
summary(Test_Data$Personal.Loan)
```

Train the model using the training set.
```{r}
set.seed(123)
UBdata_normalized$Personal.Loan<- as.factor(UBdata_normalized$Personal.Loan)
model<-train(Personal.Loan~Age+Experience+Income+Family+CCAvg+Mortgage+Securities.Account+CD.Account+Online+CreditCard+Education.1+Education.2+Education.3, data = UBdata_normalized, method = "knn")
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

